Quality Assessment of OpenStreetMap Footpath Data

### 2020-08-25 report/report.tex
English spell check

As a secondary study, the positional accuracy of the OSM dataset was calculated, since the positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
Any geodataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
\end{abstract}

@@

Som en andra del av studien har felmarginalen för OSM-datamängden också tagits fram, eftersom felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
Varje geodatamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmedoter som har använts i datans insamling.
I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
\end{abstract}
\end{otherlanguage}

@@

Virtual cities has since long been a popular research topic in the field of computer visualization.
The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modeling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
Traditionally, modeling virtual cities is a labor-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
This type of task is commonly done by professional 3D graphical artists and designers.
However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvement of a 3D artist.
\textbf{Procedural models} typically involve a machine which algorithmically composes together or alters human-generated 3D assets.
Sometimes a degree of computer-generated randomness is induced into the generative algoritm in order to create a larger and less predictable generative space of 3D models that can be created from the same algorithm.

@@

To limit visual quirks such as mesh self intersection, it is often in the designer's best interest to have a stable procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM.
The OSM project is a freely available wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
The OSM world dataset is being consistently updated by volunteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
There are already a few such services that use OSM data to generate 3D geodata visualizations for media and video game applications.
A few prominent examples are Mantle
\footnote{Mantle homepage. Accessed 17-05-2020. https://www.mantle.tech/}
and Mapbox
\footnote{Mapbox homepage. Accessed 17-05-2020. https://www.mapbox.com/}.
Such services are however usually limited in the accuracy and type of data that they can provide.
While they can provide both building footprints (the polygon that defines the building's extents) and road networks, they usually provide road data with no respect to overlap between features.

@@

Again, note that feature overlap is defined here as geometry errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two subquestions. The primary subquestion is:

@@

This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard widths according to swedish regulation.
It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
It has also identified how many of those areas can in theory be corrected by simple translation of the road vertices, without propagation of collision to other features.
Here, features are defined as the geometrical components that comprise the OSM map.
For the purpose of this report, features will refer mainly to roads and building footprints.
For the overlap study, see sections 3.1.2 for a method overview, 3.3 for a detailed walk through of the method, and 4.1.1 for the results.

The secondary subquestion is:

@@

The positional accuracy will be necessary to carry out the investigation of the second subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building dataset with known accuracy. See section 3.4.2 for an overview of the SLU dataset and how it is obtained.
The positional accuracy will be needed to assess where road overlap occurs and whether the collisions can easily be corrected.
For the positional accuracy study, see sections 3.1.1 for a method overview, 3.2 for a walk through of the method, and 4.1.2 for the results.

@@

The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighborhoods, giving the city grid its unique aestethic character and defining possible building arrangements.
Thus, for any procedural city reproduction model it will be necessary to have a robust procedure for extracting road geodata and generating the 3D representation of it without compromising the data integrity. 
So far there has been a lot of research into this area, but \textbf{(Stojanovski, 2019)} claims that most of these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularly due to the interest in self-propelled cars.
The study of energy efficient mobility such as walking and cycling is only in its emerging phase.

@@

\textbf{(Müller et al., 2006)} present how procedural model generation is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars.
Shape grammars often work from a data bank of 3D models and a set of hierarchical rules for how these can be pieced together.
This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
Using this procedure, a procedural model can structure the façade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
The maximum accuracy that can be obtained in these finer details is largely dependent on the availability of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.

@@

The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al., 2014)} in their study of the quality of the OSM Munich dataslice in 2014.
First a feature map was found, which mapped individual buildings between the OSM dataset and a reference dataset.
A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm by \textbf{(Douglas and Peucker, 1973)}.
Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.

@@

Once the positional error estimate was obtained from the precision study, it was used to assess the integrity of the OSM Stockholm road network.
This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making adjustments within the found margin of error.
For the purpose of the overlap study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
After the width assignment, a program iterated over all features and calculated how many of them that show overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
Finally, the program calculated how many of these colliding features that can be corrected by making adjustments to the feature vertices, away from the collision point, that are within the positional accuracy of the OSM data, as obtained by the geodata precision study.
The motivation behind using the positional accuracy as an upper bound for correction is to tread the positional accuracy as a maximum resolution of the map.
Making translation withing the map resolution will preserve the logical consistency to OSM source data, and preserve aestetic characteristics of the map, both between features and in the map as a whole

@@

Once a feature map has been obtained, the next task will be to find a matching vertex set among the vertices of the corresponding building footprint polygons.
Provided that the corresponding building footprints are true representations of the real life building, these pairs of matching verties must match the same part of the building.
The most common case is that a vertex represents a corner in the footprint polygon, and in these cases it must hold that two matching vertices represent the same corner.
The central problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship, since footprints from different datasets may be formed at a different level of detail.
Possible causes are that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the two polygons.
Thus, simply comparing the closest vertices between the two building representations may not be valid, since those vertices may not represent the same part of the true building footprint.
To avoid this effect, key points were extracted using the Douglas-Peucker algorithm, in order to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favour of creating polygons with reduced vertex counts.
A useful property of the Douglas-Peucker algorithm is that two highly similar polygons of different vertex counts will be reduced to resulting polygons with similar vertex counts.
Thus, one can be much more certain that a closest vertex pair is indeed corresponding to the same feature of the building.

@@

The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI).
OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
The driving force behind the availability of VGI is, according to \textbf{(Goetz and Zipf, 2012)}, the availability of high resolution satellite imagery to private individuals.
They claim that particularly the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.

@@

It was found in the feature overlap study that the individual cases of overlap could be classified into types, where all collisions of the same type could in theory be corrected by the same algoritm.
Note however that since this study focused exclusively on the Stockholm area, and the feature collisions detected in Stockholm could be of types that are characteristic to Stockholm in particular.
There could very well be additional types of collisions that are exclusive to cities with other unique city grid aestethics, which would not have been considered by this particular project.

@@

Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section 2.3.5 of this report.
It turns out that, with the current means of data acquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
If a modeler wishes to acquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aestethic depth, but do not necessarily reflect reality.

@@

Simple, procedurally generated cities, populated with generic 3D models, could very well be enough for simple presentation in e.g. animation and video games or even urban planning software, 
In such a model however, features will only be rough representations, and rarely match their real life counterparts.
Such models will be sufficient as long as one can be satisfied with looking at rough representations of major features, and do not require additional knowledge about minor features.
It turns out however, that many studies in urban development have a highly case-by-case nature.
Studies acroos many different areas have been done before that required very different knowledge about a city's geometry, e.g. for the purposes of modeling heat development
\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
or reducing noise
\textbf{(Joon, Seo and Byung, 2011)}.

According to \textbf{(Stojanovski, 2019)}, software that strives to aid in the urban development process is commonly not sophisticated enough to fully represent the needs and workspaces of urban planners in practise.
Stojanovski goes on to claim that a lot of research in this field requires detailed information that goes beyond simply knowing the geometry of the city.
Additionally, such research also typically requires additional logical data structures that a OSM cannot note serve.
Urban architects often work with morphological structures where a single feature, such as a city neighborhood, can be examined from many different levels of perspective.
A neighborhood can be analyzed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
Similarly a road network can be analyzed by looking at it either at its connectivity by viewing it a a simple 2D graph, but it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets. 
The very unfortunate fact is that much of that morphological information is not present in the OSM dataset.
Furthermore, take the fact that even a single road can be comprised of many different sections with different purposes.
A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
This type of representation is known as a multimodal structures, due to the possibility of incorporating multiple modes of transportation, but it may even describe subtilties like building facades and distances between curb line and building lot.
OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of way finding to create logical routes for different types of traffic.
For urban planners however, requiring an exact representation of some specification, this complementary data will have to be obtained from another dataset.
It may also be necessary for a researcher to implement their own data structures on top of the existing city model, with will increase the complexity of the problem even further.
The case-by-case nature of urban development projects also means that it is unlikely that a single piece of software can cater to the very differing needs of data and data structures for variuos projects.

@@

\section*{Acknowledgements}
The author would like to thank KTH associate professor Christopher Peters for supervising this master's thesis project, as well as KTH postdoctoral researcher Todor Stojanovski for expert consultation in the field of urban development.

>>>>

As a secondary study, the positional accuracy of the OSM dataset was calculated, since the positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
Any geographic dataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
\end{abstract}

@@

Som en andra del av studien har felmarginalen för OSM-datamängden också tagits fram, eftersom felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
Varje geografisk datamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmetoder som har använts i datans insamling.
I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
\end{abstract}
\end{otherlanguage}

@@

Virtual cities has since long been a popular research topic in the field of computer visualization.
The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modeling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
Traditionally, modeling virtual cities is a labor-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
This type of task is commonly done by professional 3D graphical artists and designers.
However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvement of a 3D artist.
\textbf{Procedural models} typically involve a machine which algorithmically composes together or alters human-generated 3D assets.
Sometimes a degree of computer-generated randomness is induced into the generative algorithm in order to create a larger and less predictable generative space of 3D models that can be created from the same algorithm.

@@

To limit visual quirks such as mesh self intersection, it is often in the designer's best interest to have a stable procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based map services such as OSM.
The OSM project is a freely available wiki-like geographic database based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
The OSM world dataset is being consistently updated by volunteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
There are already a few such services that use OSM data to generate 3D geodata visualizations for media and video game applications.
A few prominent examples are Mantle
\footnote{Mantle homepage. Accessed 17-05-2020. https://www.mantle.tech/}
and Mapbox
\footnote{Mapbox homepage. Accessed 17-05-2020. https://www.mapbox.com/}.
Such services are however usually limited in the accuracy and type of data that they can provide.
While they can provide both building footprints (the polygon that defines the building's extents) and road networks, they usually provide road data with no respect to overlap between features.

@@

Again, note that feature overlap is defined here as geometry errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two sub-questions. The primary sub-question is:

@@

This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard widths according to Swedish regulation.
It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
It has also identified how many of those areas can in theory be corrected by simple translation of the road vertices, without propagation of collision to other features.
Here, features are defined as the geometrical components that comprise the OSM map.
For the purpose of this report, features will refer mainly to roads and building footprints.
For the overlap study, see sections 3.1.2 for a method overview, 3.3 for a detailed walk through of the method, and 4.1.1 for the results.

The secondary sub-question is:

@@

The positional accuracy will be necessary to carry out the investigation of the second sub-question: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building dataset with known accuracy. See section 3.4.2 for an overview of the SLU dataset and how it is obtained.
The positional accuracy will be needed to assess where road overlap occurs and whether the collisions can easily be corrected.
For the positional accuracy study, see sections 3.1.1 for a method overview, 3.2 for a walk through of the method, and 4.1.2 for the results.

@@

The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighborhoods, giving the city grid its unique aesthetic character and defining possible building arrangements.
Thus, for any procedural city reproduction model it will be necessary to have a robust procedure for extracting road geodata and generating the 3D representation of it without compromising the data integrity. 
So far there has been a lot of research into this area, but \textbf{(Stojanovski, 2019)} claims that most of these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularly due to the interest in self-propelled cars.
The study of energy efficient mobility such as walking and cycling is only in its emerging phase.

@@

\textbf{(Müller et al., 2006)} present how procedural model generation is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars.
Shape grammars often work from a data bank of 3D models and a set of hierarchical rules for how these can be pieced together.
This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
Using this procedure, a procedural model can structure the facade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
The maximum accuracy that can be obtained in these finer details is largely dependent on the availability of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.

@@

The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al., 2014)} in their study of the quality of the OSM Munich data slice in 2014.
First a feature map was found, which mapped individual buildings between the OSM dataset and a reference dataset.
A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm by \textbf{(Douglas and Peucker, 1973)}.
Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.

@@

Once the positional error estimate was obtained from the precision study, it was used to assess the integrity of the OSM Stockholm road network.
This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making adjustments within the found margin of error.
For the purpose of the overlap study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
After the width assignment, a program iterated over all features and calculated how many of them that show overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
Finally, the program calculated how many of these colliding features that can be corrected by making adjustments to the feature vertices, away from the collision point, that are within the positional accuracy of the OSM data, as obtained by the geodata precision study.
The motivation behind using the positional accuracy as an upper bound for correction is to tread the positional accuracy as a maximum resolution of the map.
Making translation withing the map resolution will preserve the logical consistency to OSM source data, and preserve aesthetic characteristics of the map, both between features and in the map as a whole

@@

Once a feature map has been obtained, the next task will be to find a matching vertex set among the vertices of the corresponding building footprint polygons.
Provided that the corresponding building footprints are true representations of the real life building, these pairs of matching vertices must match the same part of the building.
The most common case is that a vertex represents a corner in the footprint polygon, and in these cases it must hold that two matching vertices represent the same corner.
The central problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship, since footprints from different datasets may be formed at a different level of detail.
Possible causes are that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the two polygons.
Thus, simply comparing the closest vertices between the two building representations may not be valid, since those vertices may not represent the same part of the true building footprint.
To avoid this effect, key points were extracted using the Douglas-Peucker algorithm, in order to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favor of creating polygons with reduced vertex counts.
A useful property of the Douglas-Peucker algorithm is that two highly similar polygons of different vertex counts will be reduced to resulting polygons with similar vertex counts.
Thus, one can be much more certain that a closest vertex pair is indeed corresponding to the same feature of the building.

@@

The OSM project is a wiki-like geographic database based on Volunteered Geographical Information (VGI).
OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
The driving force behind the availability of VGI is, according to \textbf{(Goetz and Zipf, 2012)}, the availability of high resolution satellite imagery to private individuals.
They claim that particularly the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.

@@

It was found in the feature overlap study that the individual cases of overlap could be classified into types, where all collisions of the same type could in theory be corrected by the same algorithm.
Note however that since this study focused exclusively on the Stockholm area, and the feature collisions detected in Stockholm could be of types that are characteristic to Stockholm in particular.
There could very well be additional types of collisions that are exclusive to cities with other unique city grid aesthetics, which would not have been considered by this particular project.

@@

Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section 2.3.5 of this report.
It turns out that, with the current means of data acquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
If a modeler wishes to acquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aesthetic depth, but do not necessarily reflect reality.

@@

Simple, procedurally generated cities, populated with generic 3D models, could very well be enough for simple presentation in e.g. animation and video games or even urban planning software, 
In such a model however, features will only be rough representations, and rarely match their real life counterparts.
Such models will be sufficient as long as one can be satisfied with looking at rough representations of major features, and do not require additional knowledge about minor features.
It turns out however, that many studies in urban development have a highly case-by-case nature.
Studies across many different areas have been done before that required very different knowledge about a city's geometry, e.g. for the purposes of modeling heat development
\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
or reducing noise
\textbf{(Joon, Seo and Byung, 2011)}.

According to \textbf{(Stojanovski, 2019)}, software that strives to aid in the urban development process is commonly not sophisticated enough to fully represent the needs and workspaces of urban planners in practice.
Stojanovski goes on to claim that a lot of research in this field requires detailed information that goes beyond simply knowing the geometry of the city.
Additionally, such research also typically requires additional logical data structures that a OSM cannot note serve.
Urban architects often work with morphological structures where a single feature, such as a city neighborhood, can be examined from many different levels of perspective.
A neighborhood can be analyzed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
Similarly a road network can be analyzed by looking at it either at its connectivity by viewing it a a simple 2D graph, but it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets. 
The very unfortunate fact is that much of that morphological information is not present in the OSM dataset.
Furthermore, take the fact that even a single road can be comprised of many different sections with different purposes.
A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
This type of representation is known as a multimodal structures, due to the possibility of incorporating multiple modes of transportation, but it may even describe subtilties like building facades and distances between curb line and building lot.
OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of way finding to create logical routes for different types of traffic.
For urban planners however, requiring an exact representation of some specification, this complementary data will have to be obtained from another dataset.
It may also be necessary for a researcher to implement their own data structures on top of the existing city model, with will increase the complexity of the problem even further.
The case-by-case nature of urban development projects also means that it is unlikely that a single piece of software can cater to the very differing needs of data and data structures for various projects.

@@

\section*{Acknowledgments}
The author would like to thank KTH associate professor Christopher Peters for supervising this master's thesis project, as well as KTH postdoctoral researcher Todor Stojanovski for expert consultation in the field of urban development.
